### 2020-05-13 writeup/proof_transfer_ex.tex

\textbf{Defining the multi-task learning estimator.}
Using a validation set that is sub-sampled from the original training dataset, we get a validation loss as follows
\begin{align}
		\val(\hat{B}; w_1, w_2)
	= & n_1 \cdot \bignorm{\Sigma_1^{1/2}((\frac{w_1}{w_2})^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_2^{\top}X_2(\beta_s - \frac{w_1}{w_2}\beta_t)}^2 \nonumber \\
		&+ n_1 \sigma^2 \cdot \bigtr{((\frac{w_1}{w_2})^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_1} \nonumber \\
		&+ n_2 \cdot w^2\bignorm{\Sigma_2^{1/2}((\frac{w_1}{w_2})^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_1^{\top}X_1(\beta_s - \frac{w_1}{w_2}\beta_t)}^2 \nonumber \\
		&+ n_2 \cdot \sigma^2 \cdot \bigtr{((\frac{w_1}{w_2})^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2} \label{eq_val_mtl}
\end{align}
Let $\hat{w_1}/\hat{w_2}$ be the global minimizer of $\val(\hat{B}; w_1, w_2)$.
We will define the multi-task learning estimator for the target task as
	\[ \hat{\beta}_t^{\MTL} = \hat{w}_{2}\hat{B}(\hat{w}_1, \hat{w}_2). \]

@@

\begin{lemma}\label{lem_hat_v}
	Recall that $\hat{v} = \hat{w_1} / \hat{w_2}$.
	In the setting of Proposition \ref{prop_dist_transition}, we have that
	\[ \abs{\hat{v} - 1} \lesssim \frac 1 p. \]
\end{lemma}

>>>>

\textbf{Defining the multi-task learning estimator.}
Using a validation set that is sub-sampled from the original training dataset, we get a validation loss as follows
\begin{align}
		\val(\hat{B}; w_1, w_2)
	= & n_1 \cdot \bignorm{\Sigma_1^{1/2}((\frac{w_1}{w_2})^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_2^{\top}X_2(\beta_s - \frac{w_1}{w_2}\beta_t)}^2 \nonumber \\
		&+ n_1 \sigma^2 \cdot \bigtr{((\frac{w_1}{w_2})^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_1} \nonumber \\
		&+ n_2 \cdot w^2\bignorm{\Sigma_2^{1/2}((\frac{w_1}{w_2})^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}X_1^{\top}X_1(\beta_s - \frac{w_1}{w_2}\beta_t)}^2 \nonumber \\
		&+ n_2 \cdot \sigma^2 \cdot \bigtr{((\frac{w_1}{w_2})^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2}. \label{eq_val_mtl}
\end{align}
Let $\hat{w_1}/\hat{w_2}$ be the global minimizer of $\val(\hat{B}; w_1, w_2)$.
We will define the multi-task learning estimator for the target task as
	\[ \hat{\beta}_t^{\MTL} = \hat{w}_{2}\hat{B}(\hat{w}_1, \hat{w}_2). \]

@@

We denote
\begin{align}
	\val(w) &= n_1\left[d^2 +\left( w-1\right)^2\kappa^2\right]\cdot \tr\left[( w^2X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_2^{\top}X_2)^2\right] \nonumber \\
	& + n_2w^2\left[d^2 +\left( w-1\right)^2\kappa^2\right]\cdot \tr\left[( w^2X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \nonumber\\
			& + (n_1+n_2)\sigma^2 \cdot \bigtr{(w^2X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }. \nonumber
\end{align}
Under the setting of Proposition \ref{prop_model_shift_tight}, using concentration of random vector with i.i.d. Gaussian entries, Lemma \ref{largedeviation}, we have that
$$\val(\hat{B}; w_1, w_2)= \val(w) \left( 1+\OO(p^{-1/2})\right)\quad \text{with probability }1-\oo(1).$$
Thus it suffices to study the behavior of $\val (w)$.
For the minimizer $\hat w$ of $\val(w)$, we have a similar result as in Proposition \ref{thm_cov_shift}.

\begin{lemma}\label{lem_hat_v}
Suppose the assumptions of Proposition \ref{prop_model_shift_tight} hold. Assume that $ \kappa^2 \sim pd^2 \sim \sigma^2$ are of the same order. Then we have that the optimal ratio for $\val(w)$ satisfies
	$$|\hat w -1|=\OO(p^{-1}).$$%is $\hat{w} = 1 \pm \bigo{\frac 1 {n_1+n_2}}$ \todo{(figure out the constants)}
\end{lemma}
\begin{proof}
The proof is also similar to the one for Proposition \ref{thm_cov_shift}. First it is easy to observe that $\val(w)\le \val(-w)$ for $w\ge 0$. Hence it suffices to consider the $w\ge 0$ case.

We first consider the case $w\ge 1$. We write
\begin{align}
	\val(w) &= n_1\left[\frac{d^2}{w^4} +\frac{\left( w-1\right)^2}{w^4}\kappa^2\right]\cdot \tr\left[( X_1^{\top}X_1 +w^{-2}X_2^{\top}X_2)^{-2} (X_2^{\top}X_2)^2\right] \nonumber \\
	& + n_2\left[\frac{d^2}{w^2} +\frac{\left( w-1\right)^2}{w^2}\kappa^2\right]\cdot \tr\left[( X_1^{\top}X_1 +w^{-2}X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \nonumber\\
			& + \sigma^2(n_1+n_2) \cdot \bigtr{(w^{2}X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} } .\nonumber%+  \sigma^2n_2 \cdot \bigtr{(X_1^{\top}X_1  + w^{-2}X_2^{\top}X_2)^{-1} }. \nonumber
\end{align}

Taking derivative of $\val(w)$ with respect to $w$, we obtain that
\begin{align*}
\val'(w) \ge &~ n_1\left[ \frac{2(w-1)(2-w)}{w^5}\kappa^2 - \frac{4d^2}{w^5}\right] \tr\left[( X_1^{\top}X_1 +w^{-2}X_2^{\top}X_2)^{-2} (X_2^{\top}X_2)^2\right]   \\
+&~ n_2\left[\frac{2\left( w-1\right)}{w^3}\kappa^2 - \frac{2d^2}{w^3} \right]\cdot \tr\left[( X_1^{\top}X_1 +w^{-2}X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \nonumber\\
		-&~ 2(n_1+n_2) \frac{\sigma^2}{w^3} \cdot \bigtr{(X_1^{\top}X_1 + w^{-2}X_2^{\top}X_2)^{-2} X_1^\top X_1  } = n_1 \bigtr{(X_1^{\top}X_1 + w^{-2}X_2^{\top}X_2)^{-2} \cal A },
\end{align*}
where the matrix $\cal A$ is
\begin{align*}
\cal A :&= \left[ \frac{2(w-1)(2-w)}{w^5}\kappa^2 - \frac{4d^2}{w^5}\right](X_2^{\top}X_2)^2 + \frac{n_2}{n_1} \left[\frac{2\left( w-1\right)}{w^3}\kappa^2 - \frac{2d^2}{w^3} \right](X_1^{\top}X_1)^2. - 2\frac{n_1+n_2}{n_1}\frac{\sigma^2}{w^3}X_1^\top X_1
\end{align*}
Using the estimate \eqref{eq_isometric}, we get that $\cal A$ is lower bounded as
\begin{align*}
\cal A \succeq - \frac{4d^2}{w^5}(\sqrt{n_2}+\sqrt{p})^4 + \frac{n_2}{n_1} \left[\frac{2\left( w-1\right)}{w^3}\kappa^2 - \frac{2d^2}{w^3} \right](\sqrt{n_1}-\sqrt{p})^4 - 2\frac{n_1+n_2}{n_1}\frac{\sigma^2}{w^3}(\sqrt{n_1}+\sqrt{p})^2\succ 0,
\end{align*}
as long as
$$w> w_1:=1 +\frac{d^2}{\kappa^2}+ \frac{\sigma^2}{\kappa^2}\frac{(n_1+n_2)(\sqrt{n_1}+\sqrt{p})^2}{n_2(\sqrt{n_1}-\sqrt{p})^4}+ \frac{2d^2}{\kappa^2}\frac{n_1(\sqrt{n_2}+\sqrt{p})^4}{n_2(\sqrt{n_1}-\sqrt{p})^4}.$$
Hence $\val'(w)>0$ on $(w_1,\infty)$, i.e. $\val(w)$ is strictly increasing for $w>w_1$. Hence we must have $\hat w\le w_1$. Note that under our assumptions, we have $w_1=1+\OO(p^{-1})$.

Then we consider the case $w\le 1$. Taking derivative of $\val(w)$ with respect to $w$, we obtain that
\begin{align}
	\val'(w) &\le n_1 \left[2\left( w-1\right) \kappa^2\right]\cdot \tr\left[( w^2X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_2^{\top}X_2)^2\right] \nonumber \\
	& + n_2\left[2d^2w +2w\left( w-1\right)(2w-1)\kappa^2\right]\cdot \tr\left[( w^2X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \nonumber\\
			& = n_1 \bigtr{(w^2X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} \cal B} , \nonumber
\end{align}
where the matrix $\cal B$ is
$$\cal B= 2\left( w-1\right) \kappa^2  (X_2^{\top}X_2)^2+\frac{n_2}{n_1}\left[2d^2w +2w\left( w-1\right)(2w-1)\kappa^2\right](X_1^{\top}X_1)^2.$$
Using the estimate \eqref{eq_isometric}, we get that $\cal B$ is upper bounded as
\begin{align*}
\cal B \preceq - 2(1-w)\kappa^2(\sqrt{n_2}-\sqrt{p})^4 +2d^2 w\frac{n_2}{n_1}(\sqrt{n_1}+\sqrt{p})^4 \prec 0,
\end{align*}
as long as
$$w< w_2:=1 -   \frac{d^2}{\kappa^2}\frac{n_2(\sqrt{n_1}+\sqrt{p})^4}{n_1(\sqrt{n_2}-\sqrt{p})^4}.$$
Hence $\val'(w)<0$ on $[0,w_2)$, i.e. $\val(w)$ is strictly increasing for $w<w_2$. Hence we must have $\hat w\le w_2$. Note that under our assumptions, we have $w_2=1-\OO(p^{-1})$.
\end{proof}
